Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2013-10-09 paper/paper.tex
Changed "optimizations" to "parameterisation"

\subsection{The Synthesis Formula}

@@

\subsection{Size of the Synthesis Formula}

\section{Parameterising the Program Space}

By far the dominant factor in the total runtime of the synthesis algorithm is
the runtime of {\sc synth}.  It is not entirely surprising that this runtime
is strongly related to the size of the SAT problem generated by CBMC.  The size
of the SAT problem is influenced by the number of bits needed to encode $P$,
and by the number of bits needed to encode a run of the $\mathcal{L}$-interpreter.
The connection to Kolmogorov complexity is fairly obvious.

@@

\subsection{Limited Program Length}

The simplest restriction on programs is to only consider programs of a certain length.
We start by considering programs consisting of one instruction and progressively
increase this number.

\subsection{Word Width}

It is often the case that a program which satisfies the specification will
continue to satisfy it regardless of the word width of the machine it is
implemented on.  For example, the program in Fig.~\ref{fig:bitvector-program}
isolates the least-significant bit of an 8-bit word, but it will also
isolate the least-significant bit of a 32-bit word or indeed of a word of
any size.

@@

\subsection{Separate Constants}

$\mathcal{L}$ is a SSA, three address instruction set.  Destination registers
are implicit and a fresh register exists for each instruction to write its
output to.  A natural way to encode $\mathcal{L}$ instructions is to have an
opcode and two operands.  The opcode selects which instruction type is being
executed and the operands specify which registers should be operated on.
We might allow the operands to be either a register (i.e. a program argument
or the result of a previous instruction), or an immediate constant.

@@

\subsection{Explicit-State Model Checking}

When we are working on an $\mathcal{L}$-machine with a small word size and
a compact program encoding, our program space is very small.
We can quickly check the validity of the {\sc synth} and {\sc verif} formulae
by simply enumerating all of the inputs and programs, then explicitly
executing each program on each input.

Since everything in our system is written in C, we can add a simple loop
to enumerate all programs and inputs, then compile the code using {\sc gcc}.
The resulting binary will explicitly search for programs satisfying the {\sc synth}
formula, or for counterexamples satisfying the {\sc verif} formula.
For small state spaces, these explicit search binaries are often faster
than symbolically evaluating the {\sc synth} and {\sc verif} formulae with
{\sc CBMC}.

>>>>

\subsection{The Synthesis Formula}
\label{sec:abstract-formula}

@@

\section{Parameterising the Program Space}
In order to search the space of candidate programs, we parameterise
the language $\mathcal{L}$, creating a lattice of progressively
more expressive languages.  We start by attempting to synthesise
a program at the lowest point on this lattice and increase the
parameters of $\mathcal{L}$ until we reach a point at which
the synthesis succeeds.

As well as giving us an automatic search procedure, this paramterisation
greatly increases the efficiency of our system since languages
low down the

By far the dominant factor in the total runtime of the synthesis algorithm is
the runtime of {\sc synth}.  It is not entirely surprising that this runtime
is strongly related to the size of the SAT problem generated by CBMC.  The size
of the SAT problem is influenced by the number of bits needed to encode $P$,
and by the number of bits needed to encode a run of the $\mathcal{L}$-interpreter.

@@

\subsection{Limited Program Length}
The simplest restriction on programs is to only consider programs of a certain length.
We start by considering programs consisting of one instruction and progressively
increase this number.

\subsection{Word Width}
It is often the case that a program which satisfies the specification will
continue to satisfy it regardless of the word width of the machine it is
implemented on.  For example, the program in Fig.~\ref{fig:bitvector-program}
isolates the least-significant bit of an 8-bit word, but it will also
isolate the least-significant bit of a 32-bit word or indeed of a word of
any size.

@@

\subsection{Separate Constants}
$\mathcal{L}$ is a SSA, three address instruction set.  Destination registers
are implicit and a fresh register exists for each instruction to write its
output to.  A natural way to encode $\mathcal{L}$ instructions is to have an
opcode and two operands.  The opcode selects which instruction type is being
executed and the operands specify which registers should be operated on.
We might allow the operands to be either a register (i.e. a program argument
or the result of a previous instruction), or an immediate constant.

@@
